---
title: "2-RNAseq_differential_expression"
author: "Tony Hui"
date: "March 8, 2016"
output: 
  html_document: 
    keep_md: yes
    self_contained: yes
---

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
require(tidyr)
require(knitr)
require(limma)
require(edgeR)
require(gplots)
require(pheatmap)
```

```{r, eval=FALSE}
setwd("Data_Analysis/")
```

## Load data

```{r}
rnaseq <- read.table(file="../RNASeq_data/new_data_Tony_TPM/RNAseq_new_merged_raw.txt", header = TRUE, stringsAsFactors = FALSE)

rnaseq_meta <- read.table(file = "../RNASeq_data/new_data_Tony_TPM/sailfish_file_table.txt", stringsAsFactors = FALSE)

colnames(rnaseq) <- with(rnaseq_meta, paste(V3, V4, 1:12, sep = "_"))

rnaseq_meta$samples <- with(rnaseq_meta, paste(V3, V4, 1:12, sep = "_"))
```

## Plot distribution of gExps

```{r}
rnaseq_male_Female <- rnaseq %>%
  add_rownames("gene") %>%
  select(gene, contains("vehicle")) %>%
  gather(key = sample, value = gExp, -gene) %>%
  mutate(gender = ifelse(grepl("Female", sample), "Female", "male"))

rnaseq_male_Female %>% 
  ggplot(aes(gExp+0.5, color = gender)) +
  geom_density() +
  scale_x_log10()
```

## Try Limma's `voom` function

```{r}
samples <- rnaseq_meta %>% filter(V4 == "vehicle")

limma_design_matrix <- model.matrix(~V3, samples)

rownames(limma_design_matrix) <- samples$samples

voom_rnaseq <- rnaseq %>%
  select(contains("vehicle")) %>%
  voom(design = limma_design_matrix, plot = T)

fit <- lmFit(object = voom_rnaseq, design = limma_design_matrix) %>% eBayes()

limma_results <- topTable(fit, adjust="fdr", number = Inf)
```

Pvalues are skewed to the right

```{r}
limma_results %>% 
  ggplot(aes(P.Value)) +
  geom_density()
```

### Double-check nothing funny is going on

```{r}
correlation <- cor(rnaseq %>% select(contains("vehicle")), method = "spearman")

diag(correlation) <- NA

clustering <- hclust(as.dist(1-correlation), method = "ward.D2")

require(pheatmap)
pheatmap(correlation, cluster_rows = clustering, cluster_cols = clustering, display_numbers = T, color = colorRampPalette(c("#ffffb2", "#bd0026"))(9))

plot(clustering)
```

Seems like everything is normal, although there doesn't seem to be a clear separation between male and female

## Try edgeR

```{r}
edgeR_DGElist <- rnaseq %>%
  select(contains("vehicle")) %>%
  DGEList(group = rep(c("f","m"), each = 3)) %>%
  calcNormFactors(method = "TMM") 

edgeR_DGElist_trends <- edgeR_DGElist %>%
  estimateGLMCommonDisp(limma_design_matrix, verbose=TRUE) %>%
  estimateGLMTrendedDisp(limma_design_matrix) %>%
  estimateGLMTagwiseDisp(limma_design_matrix)

plotBCV(edgeR_DGElist_trends)

plotMDS.DGEList(edgeR_DGElist_trends)

fit <- glmFit(edgeR_DGElist_trends, limma_design_matrix) %>% glmLRT(coef = 2)

edgeR_results <- topTags(fit, n = Inf) %>% as.data.frame()

edgeR_results %>% head() %>% kable("markdown")
```

Once again, right-skewed Pvalues

```{r}
qplot(edgeR_results$PValue, geom="density")
```

## Try edgeR with quasilinear fit

There is `glmFit` and `glmQLFit` - not sure the difference

Reference here: http://www.statsci.org/smyth/pubs/QLedgeRPreprint.pdf

```{r}
fitQL <- glmQLFit(edgeR_DGElist_trends, limma_design_matrix) %>% glmLRT(coef = 2)

edgeR_QL_results <- topTags(fitQL, n = Inf) %>% as.data.frame()

edgeR_QL_results %>% head() %>% kable("markdown")
```

Once again, right-skewed Pvalues

```{r}
qplot(edgeR_QL_results$PValue, geom="density")
```

## Summary of results

```{r, echo=FALSE}
paste("there are", nrow(limma_results %>% filter(adj.P.Val<0.05)), "DE genes from limma")
paste("there are", nrow(edgeR_QL_results %>% filter(FDR<0.05)), "DE genes from edgeR QL")
paste("there are", nrow(edgeR_results %>% filter(FDR<0.05)), "DE genes from edgeR")
```
```{r}
venn(list(
  edgeR_QL = edgeR_QL_results %>% filter(FDR<0.05) %>% rownames(.),
  limma = limma_results %>% filter(adj.P.Val<0.05) %>% rownames(.),
  edgeR = edgeR_results %>% filter(FDR<0.05) %>% rownames(.)
))
```

## Check for cannonical gene that should be differentially expressed

```{r}
rn6_gene <- read.table("rn6_genes.txt") %>% tbl_df() %>%
  select(gene = V1, V7) %>% 
  unique()

cannonical_gene <- c("Prl", "Xist", "Dby", "Eif2s3y", "Rps4y2", "Smcy", "Uty")

rn6_gene_interest <- rn6_gene %>%
  filter(V7 %in% cannonical_gene)

rn6_gene_interest
```

```{r}
right_join(rn6_gene, edgeR_results %>% add_rownames("gene"), by = "gene")  %>%
  filter(gene %in% rn6_gene_interest$gene) %>% kable("markdown")
```

Looks like 2 of our 4 cannonical genes are differentially expressed. Yay!