---
title: "3.1-DE_genes_femaleVSzeb"
author: "Tony"
date: "3/29/2016"
output: 
  html_document: 
    keep_md: yes
---

Find all DE genes between femaleVSzeb

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)
library(edgeR)
library(gplots)
library(pheatmap)
```

## Load data

```{r}
femaleVSmale_de_genes <- read.table("../Data_Analysis/RNAseq_result/DE_genes/glmQLFit_DE_genes.tsv", header = TRUE)

rnaseq <- read.table(file="../RNASeq_data/new_data_Tony_TPM/RNAseq_new_merged_raw.txt", header = TRUE, stringsAsFactors = FALSE) 

dim(rnaseq)

rnaseq_meta <- read.table(file = "../RNASeq_data/new_data_Tony_TPM/sailfish_file_table.txt", stringsAsFactors = FALSE)

colnames(rnaseq) <- with(rnaseq_meta, paste(V3, V4, rep(1:3, 3), sep = "_"))

rnaseq_meta$samples <- with(rnaseq_meta, paste(V3, V4, rep(1:3, 3), sep = "_"))

rn6_gene <- read.table("../Data_Analysis/rn6_genes.txt") %>% tbl_df() %>%
  select(gene = V1, V7) %>% 
  unique()
```

```{r}
edgeR_DGElist_females <- rnaseq %>%
  select(contains("female")) %>%
  DGEList(group = rep(c("f","fz"), each = 3), remove.zeros = TRUE) %>%
  .[rowSums(cpm(.) > 0.3) >= 2, , keep.lib.sizes=FALSE] %>%
  calcNormFactors() 

design_matrix <- rnaseq_meta %>% filter(V3 == "Female") %>% model.matrix(~V4, .)
rownames(design_matrix) <- edgeR_DGElist_females$samples %>% rownames()

edgeR_DGElist_females_trends <- edgeR_DGElist_females %>% 
  # estimateDisp(design_matrix)
  estimateGLMCommonDisp(design_matrix, verbose=TRUE) %>%
  estimateGLMTrendedDisp(design_matrix) %>%
  estimateGLMTagwiseDisp(design_matrix)

plotBCV(edgeR_DGElist_females_trends)
# plotMDS.DGEList(edgeR_DGElist_females_trends)
```

## Call DEGs

```{r}
fitQL <- glmQLFit(edgeR_DGElist_females_trends, design_matrix) %>% glmLRT(coef = 2)

edgeR_QL_results <- topTags(fitQL, n = Inf) %>% as.data.frame() 

edgeR_QL_results %>% head() %>% kable("markdown")
```

```{r}
qplot(edgeR_QL_results$PValue, geom="density")
```

## Check overlaps

```{r}
venn(list(
  maleVSfemale = femaleVSmale_de_genes$gene,
  zebVSfemale = rownames(subset(edgeR_QL_results, FDR < 0.05))
))
```

### Save intermediate gene list of femaleVSfemale+zeb

```{r}
edgeR_QL_results %>%
  round(3) %>%
  add_rownames("gene") %>%
  filter(FDR < 0.05) %>% 
  write.table(file = "../Data_Analysis/RNAseq_result/DE_genes/3.1-femVSfemZeb_allDE_genes.tsv", row.names = F, col.names = T, quote = F, sep = "\t")
```

## Get final gene list

Genes that are DE between female VS male or zeb

```{r}
final_dualDE_genes <- edgeR_QL_results %>%
  round(3) %>%
  add_rownames("gene") %>%
  filter(FDR < 0.05) %>%
  filter(gene %in% femaleVSmale_de_genes$gene) %>%
  mutate(logFC = logFC < 0) %>%
  select(gene, gExp_up_in_FemaleVSzeb = logFC, FDR_zeb = FDR) %>%
  inner_join(., femaleVSmale_de_genes %>% select(gene, gExp_up_in_femaleVSmale = gExp_up_in_female, FDR_male = FDR, name = V7))

x <- table(final_dualDE_genes$gExp_up_in_FemaleVSzeb == final_dualDE_genes$gExp_up_in_femaleVSmale)
```

Of the `r nrow(final_dualDE_genes)` genes DE between both conditions, `r x[2]` are in the same "direction" compared to female

### Just a quick heatmap based on these genes

```{r}
final_dualDE_genes %>% head() %>% kable("markdown")

gene_list <- final_dualDE_genes %>% .$gene

plot_heatmap <- rnaseq %>%
  select(-starts_with("Male_zeb")) %>%
  subset(rownames(.) %in% gene_list) %>%
  DGEList() %>%
  cpm()

labels <- rn6_gene %>%
  filter(gene %in% gene_list) %>%
  mutate(name = make.names(V7, unique = T)) %>%
  arrange(gene)

rownames(plot_heatmap) <- labels$name
```

Color scale represents `log10(count+1)`

```{r, fig.width=5, fig.height=14, dpi=150}
pheatmap(log(plot_heatmap+1, 10), show_rownames = T, cluster_cols = hclust(as.dist(1-cor(plot_heatmap, method = "spearman")), method = "ward.D2"), clustering_method = "ward.D2",
         color = colorRampPalette(c("#ffffb2", "#feb24c", "#bd0026"))(15))
```

Try z-score normalization within rows.

```{r, fig.width=5, fig.height=14, dpi=150}
pheatmap(log(plot_heatmap+1, 10), show_rownames = T, scale = "row", cluster_cols = hclust(as.dist(1-cor(plot_heatmap, method = "spearman")), method = "ward.D2"), clustering_method = "ward.D2",
         color = colorRampPalette(c("#4575b4", "#abd9e9", "#ffffbf", "#fdae61", "#d73027"))(15))
```

## save results

Print only the concordant genes

```{r, warning=FALSE}
final <- final_dualDE_genes %>%
  filter(gExp_up_in_FemaleVSzeb == gExp_up_in_femaleVSmale) %>%
  mutate(gExp_up_in_FemaleVSall = gExp_up_in_FemaleVSzeb) %>%
  select(gene, name, gExp_up_in_FemaleVSall)

gExp <- 
  rnaseq %>%
  # select(-starts_with("Male_zeb")) %>%
  DGEList() %>%
  calcNormFactors() %>%
  cpm() %>% 
  as.data.frame() %>%
  add_rownames("gene") %>%
  gather(key = sample, value = gExp, -gene) %>%
  mutate(group = gsub("\\_[1-9]", "", sample)) %>%
  group_by(gene, group) %>%
  summarize(mean = mean(gExp) %>% round(3)) %>%
  spread(key = group, value = mean) %>%
  ungroup() 

gExp_final <- gExp %>%
  inner_join(., final, by = "gene")

head(gExp_final) %>% kable("markdown")

gExp_final %>%
  write.table(file = "../Data_Analysis/RNAseq_result/DE_genes/3.1-femVSfemZeb_glmQLFit_DE_genes.tsv", row.names = F, col.names = T, quote = F, sep = "\t")
```
